---
title: "上海交通大学《数据科学导论》季程远：测量"
tutorial:
  id: "04-r-measurement-I"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    css: "css/tutorial.css"
runtime: shiny_prerendered
description: >
  This tutorial will cover basic descriptive statistics and
  data visualization. Companion to Chapter 3 of "Quantitative Social
  Science: An Introduction in the Tidyverse."
---

<!-- Add JavaScript code for making the exercise code larger -->

```{=html}
<script language="JavaScript" src="js/exercise-font-size.js"></script>
```
```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
library(gapminder)
library(tidyverse)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = TRUE)
tut_reptitle <- "Output Report"
data(STAR, package = "qss")
star <- STAR

classcounts <- star %>%
  group_by(classtype) %>%
  count()
```

## Handling Missing Data in R

### Small class size data


In this chapter, you'll analyze data from the STAR project, which is a four-year randomized trial on the effectiveness of small class sizes on education performance. The `star` data frame as been loaded into your space so that you can play around with it a bit.


#### Exercises
- The `tibble` package is a core part of tidyverse. It allows you to convert your traditional R data frames into tibbles, which are data frames too. But tibbles make it easier and faster to work with tidyverse. You can coerce and save a current data frame to a tibble with `as_tibble`.

```{r as_tibble-star, exercise = TRUE}

```

```{r as_tibble-star-solution}
star <- as_tibble(star)
```

```{r as_tibble-star-check}
grade_code()
```

- Use the `str` function on the `star` to see what the data looks like. You can always use `head` function to view the first six rows of the data set.

```{r str-star, exercise = TRUE}

```

```{r str-star-solution}
str(star)
```

```{r str-star-check}
grade_code()
```


- Use the `glimpse` function on the `star` to see what the data looks like. `glimpse` is particularly useful when the data set contains a long list of variables (columns), as it allows you to see every column by transposing the original data set. 

```{r glimpse-star, exercise = TRUE}

```

```{r glimpse-star-solution}
glimpse(star)
```

```{r glimpse-star-check}
grade_code()
```

- Use the `dim` function on the `star` to see what the dimensions of the data look like.

```{r dim-star, exercise = TRUE}

```

```{r dim-star-solution}
dim(star)
```

```{r dim-star-check}
grade_code()
```

- Use the `dim_desc` function on the `star` to see what the dimensions of the data look like.

```{r dim_desc-star, exercise = TRUE}

```

```{r dim_desc-star-solution}
dim_desc(star)
```

```{r dim_desc-star-check}
grade_code()
```


- Use the `summary` function on the `star` to get a sense for each variable.

```{r summary-star, exercise = TRUE}

```

```{r summary-star-solution}
summary(star)
```

```{r summary-star-check}
grade_code()
```

### Handling missing data


You probably noticed that there were some `NA` values in the data when you used the `head()` function. These are **missing values**, where the value for that unit on that variable is missing or unknown. These values pose problems when we are trying to calculate quantities of interest like means or medians because R doesn't know how to handle them.

The first tool in your toolkit for missing data is the `is.na()` function. When you pass a vector `x` to `is.na(x)`, it will return a vector of the same length where each entry is `TRUE` if the value of `x` is `NA` and `FALSE` otherwise. Using logicals, you can easily get the opposite vector `!is.na(x)` which is `TRUE` when `x` is observed and `FALSE` when `x` is missing.

#### Exercises
- Use the `is.na` and `head` functions to show whether or not the first 6 values of the `g4math` variable from the `star` data frame are missing.

```{r headna, exercise = TRUE}

```

```{r headna-solution}
head(is.na(star$g4math))
```

```{r headna-check}
grade_result(
  pass_if(~ identical(.result, head(is.na(star$g4math))))
)
```

- Use the `is.na` and `sum` functions to show how many values of the `g4math` variable are missing.


```{r sumna, exercise = TRUE}

```

```{r sumna-solution}
sum(is.na(star$g4math))
```

```{r sumna-check}
grade_code()
```


- Use the `is.na` and `mean` functions to show what proportion of the `g4math` variable is missing.


```{r meanna, exercise = TRUE}

```

```{r meanna-solution}
mean(is.na(star$g4math))
```

```{r meanna-check}
grade_code()
```



### Calculating means in the fact of missing data

Missing values makes it difficult to calculate numerical quantities of interest like the mean, median, standard deviation, or variance. Many of these function will simply return `NA` if there is a single missing value in the vector. We can instruct many function to ignore the missing values and do their calculation on just the observed data by using the `na.rm = TRUE` argument. For instance, suppose we have `x <- c(NA, 1,2,3)`, then `mean(x)` will return `NA`, but `mean(x, na.rm = TRUE)` will return `2`.

#### Exercises
- Try to calculate the `mean` of the `g4math` variable in the `star` data frame without setting `na.rm`.

```{r mean-with-na, exercise = TRUE}

```

```{r mean-with-na-solution}
mean(star$g4math)
```

```{r mean-with-na-check}
grade_code(
  correct = "This isn't that useful though!"
)
```

- Try to calculate the `mean` of the `g4math` variable when setting `na.rm = TRUE`.

```{r mean-narm, exercise = TRUE}

```

```{r mean-narm-solution}
mean(star$g4math, na.rm = TRUE)
```

```{r mean-narm-check}
grade_code()
```


### Data wrangling with tidyverse: Handling missing data

You probably noticed that there were some `NA` values in the data when you used the `str` and `glimpse` functions. These are **missing values**, where the value for that unit on that variable is missing or unknown. These values pose problems when we are trying to calculate quantities of interest like means or medians because R doesn't know how to handle them. 

The `drop_na` function looks at the data frame and removes all rows (observations) with at least one missing value. In other words, it only keeps the complete rows by deleting rows where any column (variable) is filled with a `NA` value. However, this also means that `drop_na` will cause *listwise deletion*, which gets rid of the entire row even when it contains important information in other columns. Thus, it is important to specify which columns to use to drop missing data.  

#### Exercises
- Use the `drop_na` function to drop the rows with missing values in variable `g4math`. 

```{r drop_na, exercise = TRUE}

```

```{r drop_na-solution}
star %>% drop_na(g4math)
```

```{r drop_na-check}
grade_code()
```

## Visualizing Data

### Barplots


The **barplot** is a useful way to visualize a categorical or factor variable. In this exercise, you are going to visualize the `classtype` variable from the `star` data frame, which can take on the following values:

- `1` = small class
- `2` = regular class
- `3` = regular class with aid

#### Exercises
- Use the `count` function to create a data frame of counts `classcounts` for each category of the `classtype` in the `star` data.
- Use the `geom_bar` function to plot a barplot for the `classcounts` data frame. Remeber to state `geom_bar(stat='identity')` to make sure ggpplot takes the correct input for x and y axes.
- You are encouraged to use the pipe operator.

```{r barplot, exercise = TRUE}
## creat a data frame called classcounts that has
## the counts of each category of classtype


## make a barplot with ggplot

```

```{r barplot-solution}
classcounts <- star %>%
  count(classtype) # of uniques

## make a barplot with ggplot
classcounts %>% ggplot(aes(x = classtype, y = n)) +
  geom_bar(stat='identity') 

```

```{r barplot-check}
grade_code("Awesome. The graph is looking a little unhelpful, though. Let's spruce it up.")
```

### Making the barplot readable

The default barplot usually isn't all that readable.

#### Exercises

- Use the function `scale_x_discrete(labels = c())` to name the categories of the dependent variable. Use `"Small class"` for 1, `"Regular class"` for 2, and `"Regular class with aid"` for 3. Remember to factorize the variable `classtype` in `aes()` with `factor()`, otherwise the x values will be defined as numeric rather than discrete. 
- You are encouraged to use the pipe operator.

```{r barplot-labels, exercise = TRUE}
## Use the function scale_x_discrete(labels = c()) 
## i.e., scale_x_discrete(labels = c(
##     "1" = "Small class", 
##     "2" = "" ,
##     ...))



```

```{r barplot-labels-solution}
classcounts %>% ggplot(aes(x=factor(classtype), y = n)) +
  geom_bar(stat='identity') + 
   scale_x_discrete(labels = c(
     "1" = "Small class", 
     "2" = "Regular class", 
     "3" = "Regular class with aid"))

```


- Use the function `labs` to add the lables and title to the plot. Lable the x axis `"Classroom Type"`, the y axis `"Number of students"`, and the title `"The Distribution of Students in Different Class Types"`.
- You are encouraged to use the pipe operator.


```{r barplot-lab, exercise = TRUE}
## Use the function `labs` and `title`



```

```{r barplot-lab-solution}
classcounts %>% ggplot(aes(x=factor(classtype), y = n)) +
  geom_bar(stat='identity') + 
   scale_x_discrete(labels = c(
     "1" = "Small class", 
     "2" = "Regular class", 
     "3" = "Regular class with aid")) +
  labs( x = "Classroom Type", 
        y = "Number of students", 
        title = "The Distribution of Students in Different Class Types")
```


### Histograms

For quantitative (numerical) variables, the barplot won't work because there are too many unique values. In this case, you will often use a histogram to visualize the a numerical variable.

#### Exercises
- With the `ggplot()` function, use the `geom_histogram()` geom to create a histogram for the `g4math` variable in the `star` data frame.
- You are encouraged to use the pipe operator.

```{r hist, exercise = TRUE}
## create a histogram of g4math with ggplot

```


```{r hist-solution}
## create a histogram of g4math with ggplot
star %>% ggplot(aes(x = g4math)) +
  geom_histogram()
```

```{r hist-check}
grade_code("Great job, though the graph is a bit spartan. Let's make it more readable.")
```

### Sprucing up the histogram

As with the barplot, there are several arguments you can pass to the `ggplot()` function that will improve its readability:

- `aes(y = after_stat(density))` argument in `geom_histogram` allows you to make a density plot with `ggplot`.
- `labs` allows you to add character strings that print a main title for the plot, and set the labels for the x (horizontal) and y (vertical) axes.
- `lims`: to specify the interval for the x and y axes.

#### Exercises
- Create a histogram with `ggplot` where you (a) include `aes(y = after_stat(density))` argument in `geom_histogram` to make a density plot, (b) set the y-axis to be between `0` and `0.015` using the `lims` argument, (c) include an informative x-axis label using the `labs` argument, and (d) include a title for the plot using the `labs` argument.
- Make sure to separate the arguments in function calls with commas.
- You are encouraged to use the pipe operator.


```{r hist-nice, exercise = TRUE}
## create the histogram with the specifications given in the instructions

```


```{r hist-nice-solution}
## create the histogram with the specifications given in the instructions
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = after_stat(density))) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores")

```


- Density plots can be less useful when we want to show the proportion of each bin in a histogram, and we use `y = stat(density*width)` to convert the density back to percentage. 

```{r hist-prop, exercise = TRUE}
## create the histogram that shows the proportion of each bin

```


```{r hist-prop-solution}
## create the histogram with the specifications given in the instructions
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = after_stat(density*width))) 
```


### Adding lines and text to a plot

We'll often want to add more information to a plot to make it even more readable. You can do that with geoms that add to the current plot, such as `geom_abline` and `annotate`.

- `geom_abline`: adds a line with specific slope and intercept
- `geom_vline`: adds a vertical line
- `geom_hline`: adds a horizontal line

#### Exercises
- Use the `geom_vline` function to add a vertical line at the mean of the `g4math` variable from the `star` data. By default, missing values are removed by ggplot with a warning and you can use the `na.rm = TRUE` argument to silently remove them.
- You are encouraged to use the pipe operator.

```{r hist-line, exercise = TRUE}
## add a vertical line at the mean of the variable
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = after_stat(density))) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores") +
  geom_vline(xintercept = mean( , na.rm = TRUE)) #use the dollar sign 
```

```{r hist-line-solution}
## add a vertical line at the mean of the variable with geom_vline
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = after_stat(density))) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores") +
  geom_vline(xintercept = mean(star$g4math, na.rm = TRUE))
```

```{r hist-line-check}
grade_code()
```

### Adding text to a plot

We'll sometimes want to add text to a plot to make it more informative. `annotate(geom = "text", x = 8, y = 9, label = "A")` adds a character string `A` centered at point on the (`8`, `9`) on the plot. You can use the axis labels to see where you might want to add these parts of the plot.

#### Exercise

- Use the `annotate` function to add the string `Average Score` to the plot at the point (750, 0.014).
- Make sure to separate the arguments in function calls with commas.
- You are encouraged to use the pipe operator.

```{r hist-text, exercise = TRUE}
## add the text "Average Score" at the specified location

```

```{r hist-text-solution}
## add the text "Average Score" at the specified location
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = after_stat(density))) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores") +
  geom_vline(xintercept = mean(star$g4math, na.rm = TRUE)) +
  annotate(geom = "text", x = 750, y = 0.014, label = "Average Score")
  

```

```{r hist-text-check}
grade_code()
```

## More visualization


We'll continue to analyze data from the STAR project, which is a four-year randomized trial on the effectiveness of small class sizes on education performance. The `star` data frame as been loaded into your space so that you can play around with it a bit.

### Boxplots

Boxplots are useful tools to visualize how the distribution of a continuous variable changes across levels of a categorical variable. There are several ways to specify a boxplot, but the most basic way to construct with tidyverse is as follow:

    ggplot(star, aes(y = cont.var, x = cat.var)) +
      geom_boxplot()

Here in the `ggplot` function, the first argument is name of the data frame that contains the two variables of interest, followed by `aes` to specify them on the boxplot. Most importantly, we want to specify the function `geom_boxplot` to produce a boxplot. Note that you can omit the `star` argument in the `ggplot` if you pass the data frame using the `%>%` operator.

#### Exercises

We can start by factorizing the `classtype` variable by using the `mutate` and the `as.factor` functions. Next, pass the expression to the `ggplot` function using the `%>%` operator. Include informative x-axis and y-axis labels with the  `labs` function. To fine tune the graph, we can use the `scale_x_discrete` function to give the class types more intuitive names. 

```{r boxplot, exercise = TRUE}
## factorize the variable `classtype`


## create a box plot with the characteristics specified in the instructions


```

```{r boxplot-solution}
star %>% 
  mutate(classtype = as.factor(classtype)) %>%
  ggplot(aes(x = classtype, y = g4math)) + 
  geom_boxplot() +
  labs(y = "Math Scores", x = "Class Type") +
  scale_x_discrete(labels = c("Small class", "Regular class", "Regular class with aid"))
```

### Scatter plots

Now you'll think more about to measure bivariate relationships---that is, the relationship between two variables. The `ggplot` function takes the `data` argument followed by `aes` to specify the variable-of-interests, more importantly, specify `geom_point` to create a scatter plot as such: `ggplot(data, aes(x = variable, y = variable)) + geom_point()`. You will use this plot to explore the relationship between math and reading test scores in the `star` data.

#### Exercises
- Create a scatter plot using the `ggplot` and `geom_point` function with `g4math` on the x-axis and `g4reading` on the y-axis, both from the `star` data.

```{r scatter, exercise = TRUE}
## produce a scatterplot of g4math on the x-axis and g4reading on the y-axis

```

```{r scatter-solution}
## produce a scatterplot of g4math on the x-axis and g4reading on the y-axis
ggplot(star, aes(x = g4math, y = g4reading)) +
  geom_point()
```

```{r scatter-check}
grade_code("OK, great plot! Let's make it a bit more polished.")
```


### Plotting two sets of points

Often we want to plot certain pints differently than others. For instance, maybe you want to see how the relationship between math and reading scores differs between students in small classes versus those not in small classes. To do this, we can subset the target classtypes with `filter` then specify colors of the different classtypes with `aes` within the `geom_point` function.

Before we do so, make sure to factorize the `classtype` variable for `ggplot` to create the graph.

#### Exercises

- Change the variable class of `classtype` to factor by using the `as.factor()` function
- Subset the data frame to only include Small class and Regular class for this scatter plot
- Pass the above expressions to the `ggplot` function using the `%>%` operator
- Differentiate data points from Small and Regular class by specifying `color` in the aes nested in `geom_point` 


```{r plot-points, exercise = TRUE}

```


```{r plot-points-solution}
## tidyverse
star %>% 
  mutate(classtype = as.factor(classtype)) %>%
  filter(classtype %in% c("1", "2")) %>%
  ggplot(aes(x = g4math, y = g4reading)) + 
  geom_point(aes(color = classtype)) 
  
```

```{r plot-points-check}
grade_this_code()
```

### Finalizing your scatter plot

The scatter plot is looking very good, but it could use a little bit of polish. Let's add axis labels and a title. 

#### Exercises
- To do so, take the current plot and pass it to the `labs` function to make the graph more informative.

```{r niceplot, exercise = TRUE}

```

```{r niceplot-solution}
star %>% 
  mutate(classtype = as.factor(classtype)) %>%
  filter(classtype %in% c("1", "2")) %>%
  ggplot(aes(x = g4math, y = g4reading)) +
  geom_point(aes(color = classtype)) +
  labs(x = "Fourth Grade Math Scores", 
       y = "Fourth Grade Reading Score", 
       title = "Math vs Reading")
```


## Conceptual Questions


```{r concept, echo=FALSE}
quiz(caption = "",
     question("In section 3.6 of QSS, what kind of plot are we using to visualize ideological differences between Democrats and Republicans?",
              answer("scatterplot", correct = TRUE),
              answer("bar plot"),
              answer("histogram")),
     question("To estimate the correlation between two variables x and y, we need to know the mean of each as well as the:",
              answer("maximum"),
              answer("standard deviation", correct = TRUE),
              answer("length")),
     question("What type of relationships does correlation measure?",
              answer("linear relationships", correct = TRUE),
              answer("nonlinear relationships"),
              answer("both"))
     )
```

## Correlation and scaling

To illustrate the basics of correlation, we'll focus again on the `gapminder` data which has a few continuous variables

```{r}
gapminder
```

Let's use this data to explore correlation as a summary measure of the relationship between two variables. Correlations between variable can be calculated by using the syntax `cor(x, y)` where `x` and `y` are two vectors of the same length that we want to calculate the correlation of. We can do this in two ways for a data set. First, we can use the `$` operator to extract variables from our data like so:

```{r eval = FALSE}
cor(mydata$x, mydata$y)
```

Second, we can use the pipe and summarize combination:

```{r eval = FALSE}
mydata %>%
  summarize(cor(x, y))
```

Let's try the first approach to find the correlation between `lifeExp` and `gdpPercap` in the gapminder data:




```{r cor1, exercise = TRUE}

```

```{r cor1-solution}
cor(gapminder$lifeExp, gapminder$gdpPercap)
```

```{r cor1-check}
grade_this_code()
```

Now let's try it the second way using pipe+summarize:


```{r cor2, exercise = TRUE}

```

```{r cor2-solution}
gapminder %>%
  summarize(cor(lifeExp, gdpPercap))
```

```{r cor2-check}
grade_this_code()
```

### Z-scores

Correlations are calculated using z-scores of each variable. A z-score of a variable is a transformation that puts all variables on a common scale. This transformation subtracts off the mean of the variable from each unit and then divides by the standard deviation of the variable. We could do this transformation with:

```{r eval=FALSE}
mydata %>%
  mutate(my_z = (x - mean(x)) / sd(x))
```

Let's now calculate the correlation "by hand" by creating z-scores for both `lifeExp` and `gdpPercap` and then taking the average of the product of these. 



```{r scale, exercise = TRUE}
gapminder %>%
  mutate(
    lifeExp_z = ___,
    gdpPercap_z = ___
  ) %>%
  summarize(cor = mean(___))
```

```{r scale-solution}
gapminder %>%
  mutate(
    lifeExp_z = (lifeExp - mean(lifeExp)) / sd(lifeExp),
    gdpPercap_z = (gdpPercap - mean(gdpPercap)) / sd(gdpPercap)
  ) %>%
  summarize(cor = mean(lifeExp_z * gdpPercap_z))
```

```{r scale-check}
grade_this_code()
```


## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
